---
title: "Predicting Financial behavior"
author: "PK (Kasidit) Ratanavijai"
date: "10/25/2019"
output: github_document
---

```{r 1, include=FALSE}
# install if require
if(!require(zoo)) install.packages("zoo") 
if(!require(tidyverse)) install.packages("tidyverse") 
if(!require(kableExtra)) install.packages("kableExtra")
if(!require(tidyr)) install.packages("tidyr")
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(stringr)) install.packages("stringr")
if(!require(forcats)) install.packages("forcats")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(GGally)) install.packages("GGally")
if(!require(corrgram)) install.packages("corrgram")
if(!require(fastR2)) install.packages("fastR2")
if(!require(mosaic)) install.packages("mosaic")
# Loading all libraries
library(zoo)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(tidyr)
library(stringr)
library(forcats)
library(ggplot2)
library(fastR2)
library(mosaic)
library(janitor)
library(lubridate)
library(corrgram)
library(GGally)

```

```{r 2, include=FALSE}
#load data
dataPath = "C:/Users/PK/Downloads/Statfinal"

AssignmentData<-
  read.csv(file=paste(dataPath,"datastat.csv",sep="/"),
           row.names=1,header=TRUE,sep=",")
```


# Project objectives

Predicting cycles when FED implemented easing policy (reducing rates) or tightening policy (increasing rates) and Investor's behavior towards long term and short term financial assest by using "DAILY TREASURY YIELD CURVE RATES" from year January 1981 to June 2014 total of 33 years.

The data set composed of date, rate of US Generic Government teasury  3 months yield, 6 months, 2 years, 3 years, 5 years, 10 years and 30 years, and whether FED was using easing policy (reducing rates) or tightening policy (increasing rates) at the time.


It is public data set and can be found on *treasury.gov* (https://home.treasury.gov/policy-issues/financing-the-government/interest-rate-statistics)

# Data exploration

**Data Summary**

```{r echo=FALSE}
summary(AssignmentData)
```

**Structure of the data set**

- **USGG3M** ```<numeric>``` rate of US Generic Government teasury  3 months yield
- **USGG6M** ```<numeric>``` rate of US Generic Government teasury  6 months yield
- **USGG2YR** ```<numeric>``` rate of US Generic Government teasury  2 years yield
- **USGG5YR** ```<numeric>``` rate of US Generic Government teasury  3 years yield
- **USGG10YR** ```<numeric>``` rate of US Generic Government teasury  5 years yield
- **USGG30YR** ```<numeric>``` rate of US Generic Government teasury  10 years yield
- **Output1** ```<numeric>``` sssssxxxx
- **Easing** ```<int>``` whether FED was using easing policy (reducing rates)
- **Tightening** ```<int>``` whether FED was using tightening policy at the time (increasing rates)

First 7 variables (input variables) are daily records of the US Treasury yields to maturity. The last two variables mark periods when FED implemented easing policy (reducing rates) or tightening policy (increasing rates).

**Header of the training data set**

```{r 3, echo=FALSE}
head(AssignmentData)
```

**History of Interest Rates**

```{r 4, echo=FALSE}
matplot(AssignmentData[,-c(8,9,10)],type='l',ylab="Interest Rates",
        main="History of Interest Rates",xlab="Index")
```

**History of Interest Rates and Output**

```{r 5, echo=FALSE}
matplot(AssignmentData[,-c(9,10)],type='l',ylab="Interest Rates",
        main="History of Interest Rates and Output",xlab="Index")
```

# Data preparation

**from Data exploration we can see that there are NAs in Easing and Tightening**
Make column "Tightening" equal to 1 during tightening periods and 0 otherwise.


```{r include=FALSE}
AssignmentDataLogistic<-AssignmentData[,c(1:7,10)]
All.NAs<-is.na(AssignmentData[,9])&is.na(AssignmentData[,10]) # neither tightening nor easing
noTightening<-is.na(AssignmentData[,10]) 
AssignmentDataLogistic[noTightening,'Tightening']<-0 # replace NAs with 0
cat("Before: ",dim(AssignmentDataLogistic),"\n") # before
AssignmentDataLogistic<-AssignmentDataLogistic[!All.NAs,]
AssignmentDataLogistic[c(275:284),]
```


# Modelling 

## Linear regression model

**Fit 7  regression models, in each of them input variable of interest rate is response and Output1 is the single predictor**

```{r include=FALSE}
# need to assign new name to match with Output1 name in file newPredictor else does not work
Output1 = AssignmentData$Output1

```


```{r echo=TRUE}
#Fit 7 simple regression models
mUSGG3M = lm(AssignmentData$USGG3M ~ Output1)
mUSGG6M = lm(AssignmentData$USGG6M ~ Output1)
mUSGG2YR = lm(AssignmentData$USGG2YR ~ Output1)
mUSGG3YR = lm(AssignmentData$USGG3YR ~ Output1)
mUSGG5YR = lm(AssignmentData$USGG5YR ~ Output1)
mUSGG10YR = lm(AssignmentData$USGG10YR ~ Output1)
mUSGG30YR = lm(AssignmentData$USGG30YR ~ Output1)
```

Here is significance of the estimated parameters and the model **mUSGG3M** as a whole, proportion of explained correlation.

```{r echo=FALSE}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(mUSGG3M)$sigma^2)
```

and coefficient

```{r echo=FALSE}
coef(mUSGG3M)
```

Plot the output variable together with the fitted values.

```{r 9, echo=FALSE}
matplot(AssignmentData[,1],type="l",col="blue",xaxt="n")
lines(mUSGG3M$fitted.values,col="orange")
```

Collect all slopes and intercepts in one matrix called simpleRegressionResults and keep this matrix
```{r echo=FALSE}
simpleRegressionResults =data.frame()
s =rbind (USGG3M=mUSGG3M$coefficients,USGG6M=mUSGG6M$coefficients,USGG2YR=mUSGG2YR$coefficients,USGG3Y=mUSGG3YR$coefficients,USGG3Y=mUSGG3YR$coefficients,USGG5Y=mUSGG5YR$coefficients,USGG10Y=mUSGG10YR$coefficients,USGG30Y=mUSGG30YR$coefficients)
simpleRegressionResults = data.frame(s)
names(simpleRegressionResults) = c('regrIntercepts','Slopes')
simpleRegressionResults
```


Use fitted models for newPredictor with file 'Output_for_prediction.csv'

```{r include=FALSE}
newPredictor = read.csv('Output_for_prediction.csv')
```


```{r echo=TRUE}
#Use function predict() to predict 7 interest rates for the value of newPredictor
prUSGG3M=predict(mUSGG3M,newdata=newPredictor)
prUSGG6M=predict(mUSGG6M,newdata=newPredictor)
prUSGG2YR=predict(mUSGG2YR,newdata=newPredictor)
prUSGG3YR=predict(mUSGG3YR,newdata=newPredictor)
prUSGG5YR=predict(mUSGG5YR,newdata=newPredictor)
prUSGG10YR=predict(mUSGG10YR,newdata=newPredictor)
prUSGG30YR=predict(mUSGG30YR,newdata=newPredictor)
```

**examine predicted.values**

```{r echo=FALSE}
c(`prUSGG3M`, `prUSGG6M`, `prUSGG2YR`, `prUSGG3YR`, `prUSGG5YR`, `prUSGG10YR`, `prUSGG30YR`)
```

## Logistic regression model

**Plot the data and the binary output variable representing easing (0) and tightening (1) periods.**

```{r 15, echo=FALSE}
matplot(AssignmentDataLogistic[,-8],type="l",ylab="Data and Binary Fed Mode",ylim=c(-5,20))
lines(AssignmentDataLogistic[,8]*20,col="blue",lwd=3)
legend("topright",legend=c("response"),lty=1,col=c("blue"),lwd=3)
```

Estimate logistic regression with 3M yields as predictor and Tightening as output.


```{r echo=TRUE}
LogisticModel_3M<-glm(Tightening~USGG3M,family=binomial(link=logit),AssignmentDataLogistic)
summary(LogisticModel_3M)
```

Plot the data, the response and the predicted probability of tightening.

```{r 16, echo=FALSE}
matplot(AssignmentDataLogistic[,-8],type="l",ylab="Data and Fitted Values",ylim=c(-5,25))
lines(AssignmentDataLogistic[,8]*20,col="blue",lwd=3)
lines(LogisticModel_3M$fitted.values*20,col="orange",lwd=3)
legend("topright",legend=c("prediction","response"),lty=1,col=c("orange","blue"),lwd=3)
```

Now use all inputs as predictors for logistic regression. Name the model LogisticModel_All.

```{r echo=TRUE}
LogisticModel_All = glm(Tightening~USGG3M+USGG6M+USGG2YR+USGG3YR +USGG5YR +USGG10YR +USGG30YR  ,family=binomial(link=logit),AssignmentDataLogistic)
summary(LogisticModel_All)
```

Plot the data, predicted probability of tightening and the response.

```{r 20, echo=FALSE}
matplot(AssignmentDataLogistic[,-8],type="l",ylab="Results of Logistic Regression",ylim=c(-5,20))
lines(AssignmentDataLogistic[,8]*20,col="blue",lwd=3)
lines(LogisticModel_All$fitted.values*20,col="orange",lwd=3)
legend("topright",legend=c("prediction","response"),lty=1,
       col=c("orange","blue"),lwd=3)
```

Calculate and plot log-odds and probabilities. Compare probabilities with fitted values.
```{r echo=TRUE}
# Calculate odds
Log.Odds<-predict(LogisticModel_All)     # predict log-odds
Probabilities<-1/(exp(-Log.Odds)+1)      # predict probabilities
```


```{r 26, echo=FALSE}
# Plot log-odds and response
plot(Log.Odds,type="l",ylab="Response & Log-Odds",lwd=3,col="blue",
     main="Log-odds and Response")
lines(AssignmentDataLogistic[,8],col="orange",lwd=3)
legend("topright",legend=c("Log-Odds","Response"),
       lty=1,lwd=3,col=c("blue","orange"))
```

Use logistic regression to predict probabilities of tightening for new input data.
```{r echo=TRUE}
# New Predictors
(newPredictors<-unname(unlist(AssignmentDataLogistic[1,1:7])))
```


```{r include=FALSE}
(newPredictors<-unname(unlist(AssignmentDataLogistic[1,1:7])))
newPredictors
#Make it a data frame with the same component names as in the training data:
(newPredictors<-setNames(newPredictors,colnames(AssignmentDataLogistic[1:7])))
(newPredictors<-as.data.frame(t(newPredictors)))
newPredictors
#Use the created data frame as newdata in function predict() with type response.
newPredictors
predict(LogisticModel_All,newdata=newPredictors,type='response')
#Course Assignment. Step 2.
#Read the data from StatisticalAnalysis_Course_Assignment_2_Data.rds.
data2 <- readRDS(paste(dataPath,'StatisticalAnalysis_Course_Assignment_2_Data.rds',sep = '/'))
data2[[1]]
data3 = newPredictors
data3[1:7]=data2[[1]]
```

```{r echo=TRUE}
predict(LogisticModel_All,newdata=data3,type='response')
summary(LogisticModel_All)
```

## Comparing regression models

**Compare linear regression models with Output1 as response and different combinations of predictors.**

```{r include=FALSE}
AssignmentDataRegressionComparison<-AssignmentData[,-c(9,10)]

```


Comparing Full model with null model

```{r echo=TRUE}
#complete model 
model_all = lm(Output1~USGG3M+USGG6M+USGG2YR+USGG3YR +USGG5YR +USGG10YR +USGG30YR,AssignmentDataRegressionComparison)
#null model
model_null = lm(Output1~1,AssignmentDataRegressionComparison)
```

Compare AIC Value
```{r echo=TRUE}
anova(model_all,model_null)
c(Full=AIC(model_all),Null=AIC(model_null))
```

Lets compare nested model using anove, small model (Predict output with USGG3M) with big model (Predict output with USGG3M+ other rates)
```{r echo=TRUE}
#Given US Treasury rate = USGG3M
#Let the given rate be testRate. Let additionalRate be one of the remaining rates

mUSGG3M.2 = lm(Output1 ~ USGG3M , AssignmentData)
mUSGG6M.2 = lm(Output1 ~ USGG3M+USGG6M , AssignmentData)
mUSGG2YR.2 = lm(Output1 ~ USGG3M+USGG2YR , AssignmentData)
mUSGG3YR.2 = lm(Output1 ~ USGG3M+USGG3YR , AssignmentData)
mUSGG5YR.2 = lm(Output1 ~ USGG3M+USGG5YR , AssignmentData)
mUSGG10YR.2 = lm(Output1 ~ USGG3M+USGG10YR , AssignmentData)
mUSGG30YR.2 = lm(Output1 ~ USGG3M+USGG30YR , AssignmentData)

SumSq = c(anova(mUSGG3M.2,mUSGG6M.2)$'Sum of Sq'[2],
anova(mUSGG3M.2,mUSGG2YR.2)$'Sum of Sq'[2],
anova(mUSGG3M.2,mUSGG3YR.2)$'Sum of Sq'[2],
anova(mUSGG3M.2,mUSGG5YR.2)$'Sum of Sq'[2],
anova(mUSGG3M.2,mUSGG10YR.2)$'Sum of Sq'[2],
anova(mUSGG3M.2,mUSGG30YR.2)$'Sum of Sq'[2])

AIC = c(
AIC(mUSGG6M.2),
AIC(mUSGG2YR.2),
AIC(mUSGG3YR.2),
AIC(mUSGG5YR.2),
AIC(mUSGG10YR.2),
AIC(mUSGG30YR.2))
AIC
```

Compare SumSq and AIC explain by bigger model 

```{r}
SumSq
AIC
```

## Rolling Window Analysis

Set window width and window shift parameters for rolling window.
```{r echo=TRUE}
Window.width<-20; Window.shift<-5
```


Find mean for rolling window.
```{r echo=TRUE}
# Means
all.means<-rollapply(AssignmentDataRegressionComparison,width=Window.width,
                     by=Window.shift,by.column=TRUE, mean)
head(all.means)
```
Plot the rolling means with the original 3-month rates data

```{r}
Count<-1:dim(AssignmentDataRegressionComparison)[1]
Rolling.window.matrix<-rollapply(Count,width=Window.width,by=Window.shift,by.column=FALSE,FUN=function(z) z)
```

```{r 55, echo=FALSE}
# Find middle of each window
Points.of.calculation<-Rolling.window.matrix[,10]    
Points.of.calculation[1:10]

Means.forPlot<-rep(NA,dim(AssignmentDataRegressionComparison)[1])
Means.forPlot[Points.of.calculation]<-all.means[,1]
cbind(1:25,Means.forPlot[1:25])

plot(AssignmentDataRegressionComparison[,1],type="l",col="blue",lwd=2,
     ylab="Interest Rate & Rolling Mean", main="Rolling Mean of USGG3M")
points(Means.forPlot,col="orange",pch=1)
legend("topright",legend=c("USGG3M","Rolling Mean"),col=c("blue","orange"),lwd=2)
```



Use rolling apply to find Coefficients, R.squared, P-value, and predited value for any specific date.
```{r include=FALSE}
rolling.dates = AssignmentDataRegressionComparison[1,]
for (i in 2:nrow(AssignmentDataRegressionComparison)) {
  result = AssignmentDataRegressionComparison[i,] - AssignmentDataRegressionComparison[i-1,]
  rolling.dates = rbind(rolling.dates,result)
}
rolling.dates = rolling.dates[-1,]
head(rolling.dates)
#head(round(rolling.sd,3))
rolling.dates<-rollapply(AssignmentDataRegressionComparison[-1,],
                         width=Window.width,by=Window.shift,
                         by.column=FALSE,FUN=function(z) rownames(z))

# Rolling lm coefficients
Coefficients<-rollapply(AssignmentDataRegressionComparison,
                        width=Window.width,
                        by=Window.shift,by.column=FALSE,
                        FUN=function(z) coef(lm(Output1~USGG3M+USGG5YR+USGG30YR,
                                                data=as.data.frame(z))))

rolling.dates<-rollapply(AssignmentDataRegressionComparison[,1:8],
                         width=Window.width,by=Window.shift,by.column=FALSE,
                         FUN=function(z) rownames(z))

rownames(Coefficients)<-rolling.dates[,10]
head(Coefficients)

idxDate<-match("9/22/2005",rolling.dates[,10])
cfnts = c(Coefficients[idxDate,])
cfnts
##Find R-squared values of rolling-window-fit linear models:
# R-squared
r.squared<-rollapply(AssignmentDataRegressionComparison,
                     width=Window.width,by=Window.shift,by.column=FALSE,
                     FUN=function(z) summary(lm(Output1~USGG3M+USGG5YR+USGG30YR,
                                                data=as.data.frame(z)))$r.squared)
r.squared<-data.frame(Date=rolling.dates[,10],R2=r.squared)

head(r.squared)
testDate="9/22/2005"
idxDate<-match("9/22/2005",rolling.dates[,10])
rsqrd = c(r.squared[idxDate,2])

rsqrd
#Analyze the rolling p-values of slopes. What were dates of highest p-values (most insignificant slopes)?
# P-values
# P-values
Pvalues<-rollapply(AssignmentDataRegressionComparison,width=Window.width,
                   by=Window.shift,by.column=FALSE,
                   FUN=function(z) summary(lm(Output1~USGG3M+USGG5YR+USGG30YR,                                                                              data=as.data.frame(z)))$coefficients[,4])
rownames(Pvalues)<-rolling.dates[,10]
Pvalues[1:6,]

pvls=c(Pvalues[idxDate,])
pvls
#find fitted value
#test fitted
fitted<-rollapply(AssignmentDataRegressionComparison,
                     width=Window.width,by=Window.shift,by.column=FALSE,
                     FUN=function(z) fitted(lm(Output1~USGG3M+USGG5YR+USGG30YR,
                                                data=as.data.frame(z))))
fitted
rownames(fitted) = rolling.dates[,10]
head(fitted)

prdctn = (c(fitted[idxDate,10]))
prdctn

#predictor with highest coefficients
highestSensitivity = 'USGG5YR'

#result
res<-list(Date=testDate,Coefficients=cfnts,
          P_values=pvls,R_squared=rsqrd,Prediction=prdctn,
          HighSensitivity=highestSensitivity)

```

Look at pairwise X-Y plots of regression coefficients for the 3M, 5Yr and 30Yr yields as inputs.

```{r 33}
# Pairs plot of Coefficients
pairs(Coefficients)
```

```{r}
res
```

# PCA

*Treasury yield has high correlated to each other that makes our model risky for Multicollinearity. I will use PCA to decompose factors and loadings to solve this problem and at the same time preserve relationship from each Treasury yield and use in palce of predictors*

```{r}
AssignmentDataPCA<-AssignmentData[,1:7]
dim(AssignmentDataPCA)
```

Explore the dimensionality of the set of 3M, 2Y and 5Y yields.

```{r}
# Select 3 variables. Explore dimensionality and correlation 
AssignmentData.3M_2Y_5Y<-AssignmentDataPCA[,c(1,3,5)]
ggpairs(AssignmentData.3M_2Y_5Y)
library(rgl);rgl.points(AssignmentData.3M_2Y_5Y)
```

Load data for date we want to see change

```{r}
data <- readRDS(paste(dataPath,'StatisticalAnalysis_Course_Assignment_5_Data.rds',sep = '/'))
data
Maturities<-c(.25,.5,2,3,5,10,30)
```

See importance of factors.

```{r}
Eigen.Decomposition = princomp(AssignmentDataPCA)
Eigen.Decomposition
```

Find loadings and factors from principal component

```{r}
#find loadings
Loadings = Eigen.Decomposition$loadings
#find factors 
Factors = Eigen.Decomposition$scores
```

Change sign and plot each factors
```{r 88, echo=FALSE}
Loadings[,1]<--Loadings[,1]
Factors[,1]<--Factors[,1]

matplot(Factors,type="l",col=c("black","blue","orange"),lty=1,lwd=3,main="Factors After Sign Change")
legend("topright",legend=c("F1","F2","F3"),col=c("black","blue","orange"),lty=1)

matplot(Maturities,Loadings,type="l",lty=1,col=c("black","blue","orange"),lwd=3,main="Loadings After Sign Change")
legend("bottomright",legend=c("L1","L2","L3"),col=c("black","blue","orange"),lty=1)

```

Analyze the adjustments that each factor makes to the term curve.

```{r 46, echo=TRUE}
OldCurve<-AssignmentDataPCA[135,]
NewCurve<-AssignmentDataPCA[136,]
CurveChange<-NewCurve-OldCurve
FactorsChange<-Factors[136,]-Factors[135,]
ModelCurveAdjustment.1Factor<-OldCurve+t(Loadings[,1])*FactorsChange[1]
ModelCurveAdjustment.2Factors<-OldCurve+t(Loadings[,1])*FactorsChange[1]+t(Loadings[,2])*FactorsChange[2]
ModelCurveAdjustment.3Factors<-OldCurve+t(Loadings[,1])*FactorsChange[1]+t(Loadings[,2])*FactorsChange[2]+
  t(Loadings[,3])*FactorsChange[3]
matplot(Maturities,
        t(rbind(OldCurve,NewCurve,ModelCurveAdjustment.1Factor,ModelCurveAdjustment.2Factors,
                ModelCurveAdjustment.3Factors)),
        type="l",lty=c(1,1,2,2,2),col=c("black","red","green","blue","magenta"),lwd=3,ylab="Curve Adjustment")
legend(x="topright",c("Old Curve","New Curve","1-Factor Adj.","2-Factor Adj.",
                      "3-Factor Adj."),lty=c(1,1,2,2,2),lwd=3,col=c("black","red","green","blue","magenta"))

```

